How Kids Code and How We Know:\\An Exploratory Study on the Scratch Repository

### 2016-04-22 Paper/paper.tex
further changes

\begin{figure}
	\begin{center}
		\includegraphics[width=2cm]{fig/bump.png}
		\caption{The script for the Piano sprite in Figure \ref{fig:ui}, causing a sound on receiving the `bump' signal.}
		\label{fig:bump}
	\end{center}
\end{figure}

@@

\subsubsection{Signals}
Events within Scratch can be user generated too: users can broadcast a message, for example when two sprites touch each other, like in Figure \ref{fig:ui}. All other sprites can then react by using the \texttt{when I receive} Hat block. In Figure \ref{fig:ui}, Sprite1 broadcasts `bump' when the cat touches the Piano. The Piano reacts on that by making a sound, see Figure \ref{fig:bump}.

\subsubsection{Procedures}
Scratch users can define their own blocks, with a pop-up screen displayed in on the left hand side of Figure \ref{fig:twoBlocks}. As shown in this figure, Scratch users can enter the name of the sprite and define input parameters. The block in Figure \ref{fig:twoBlocks} uses two parameter, one textual (string1) and one numerical (number1). A user-created block is programmed using ordinary Scratch blocks as shown on the right hand side of Figure \ref{fig:twoBlocks}.

\begin{figure}
	\begin{center}
		\includegraphics[width=8cm]{fig/twoBlocks.png}
		\caption{The user interface for creating (left) and defining (right) user-defined blocks.}
		\label{fig:twoBlocks}
	\end{center}
\end{figure}

@@

Apart from whole scripts we also examined cases with only the first block being different. In first block is commonly the event block and, by separating this one from the subsequent functionality blocks, we could examine if programmers assign the same functionality to handle different types of events. Cloned functionality blocks are found to be rare: without considering the first block, only 2,243 additional clones were found in 920 projects.

The second smell that we examine is the dead code smell. Already in the previous section, Scratch projects were found to have two types of dead code: defined procedures that are not called (5.06\% of the defined procedures in 2,079 projects) and unmatched broadcast-receive messages (in 7.99\% of the projects). In this section we examine an additional case: scripts that are never invoked due to the lack of a starting \texttt{when <trigger>} block from the Scratch Events or Control category, like in the example in Figure \ref{fig:Unconnected_clones}, or scripts that are comprised of only a \texttt{when <trigger>} block without any functionality. A total of 322,475 scripts like that were found in 56,890 (24.36\% of the total) projects. The majority of those scripts (86.6\%) are scripts missing the starting block. Examining the size of those dead scripts, 72.34\% are composed of a single block. As shown in diagram \ref{fig:deadcode}, however, considerably large dead scripts exist; 2,358 of those scripts originating from 1,553 different projects have more than 30 blocks and up to 2,610\footnote{\todo{link}}.

@@

\subsubsection{Sharing of Scripts and Procedures}
Investigating the use of clones between programs, we observe that there are 1700 scripts that are used in multiple programs, sometimes as often as in 1600 different programs. This seems to indicate that there are common patterns in Scratch programs. This again leads us to the idea that it might be very beneficial to Scratch programmers if they could not only share their programs, but also share some of their functionality, for others to use, like a library. An example of such a library could be: functions for platforming games, including the movement of a player, collision detection and the implementation of `lives'. This might empower new Scratch users to get started quicker.

>>>>

%\begin{figure}
%	\begin{center}
%		\includegraphics[width=2cm]{fig/bump.png}
%		\caption{The script for the Piano sprite in Figure \ref{fig:ui}, causing a sound on receiving the `bump' signal.}
%		\label{fig:bump}
%	\end{center}
%\end{figure}

@@

\subsubsection{Signals}
Events within Scratch can be user generated too: users can broadcast a message, for example when two sprites touch each other, like in Figure \ref{fig:ui}. All other sprites can then react by using the \texttt{when I receive} Hat block. In Figure \ref{fig:ui}, Sprite1 broadcasts `bump' when the cat touches the Piano.

\subsubsection{Custom Blocks}
Scratch users can define their own blocks, which users can name themselves, called custom blocks. To align with other papers on this topic \todo{Fenia, a reference?} we will refer to them as `procedures' in the following. Procedures can have input parameters of type string, number, and boolean. When a user defines a procedure, a new Hat block called \texttt{define} appears, which users can fill with the implementation of their block.

%\begin{figure}
%	\begin{center}
%		\includegraphics[width=8cm]{fig/twoBlocks.png}
%		\caption{The user interface for creating (left) and defining (right) user-defined blocks.}
%		\label{fig:twoBlocks}
%	\end{center}
%\end{figure}

@@

Apart from whole scripts we also examined cases where scripts differed only in the first (Hat) block.This way we examine if Scratch programmers assign the same functionality to handle different types of events. Cloned functionality blocks are found to be rare: without considering the first block, only 2,243 additional clones were found in 920 projects.

The second smell that we examine is the dead code smell. Already in the previous section, Scratch projects were found to have two types of dead code: defined procedures that are not called (5.06\% of the defined procedures in 2,079 projects) and unmatched broadcast-receive messages (in 7.99\% of the projects). In this section we examine an additional case: scripts that are never invoked due to the lack of a starting \texttt{when <trigger>} block from the Scratch Events or Control category, like in the example in Figure \ref{fig:Unconnected_clones}, or scripts that are comprised of only a \texttt{when <trigger>} block without any functionality. A total of 322,475 scripts like that were found in 56,890 (24.36\% of the total) projects. The majority of those scripts (86.6\%) are scripts missing the starting block. Examining the size of those dead scripts, 72.34\% are composed of a single block. As shown in Figure \ref{fig:deadcode}, however, considerably large dead scripts exist; 2,358 of those scripts originating from 1,553 different projects have more than 30 blocks and up to 2,610\footnote{\todo{link}}.

@@

\subsubsection{Sharing of Scripts and Procedures}
Investigating the use of clones between programs, we observe that there are 1700 scripts that are used in multiple programs, sometimes as often as in 1600 different programs. This seems to indicate that there are common patterns in Scratch programs, which means it might be very beneficial to Scratch programmers if they could not only share their programs, but also share some of their functionality, for others to use, like a library. An example of such a library could be: functions for platforming games, including the movement of a player, collision detection and the implementation of `lives'. This might empower new Scratch users to get started quicker.
